Understanding Factors Contributing to a Successful Himalayan Expedition

### 2024-04-18 paper/paper.qmd
styler ran on qmd

plot_bar(analysis_data, "sex", "Sex")
plot_bar(analysis_data, "solo", "Solo Status")
plot_bar(analysis_data, "age_range", "Age")
plot_bar(analysis_data, "seasons", "Season")
plot_bar(analysis_data, "height_range", "Height of the Peak")

```

@@

  p <- ggplot(plot_data, aes(x = x_var, y = proportion, fill = factor(success))) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red"), 
                      labels = c("Success", "Died"),
                      name = "Outcome") +
    labs(x = x_label, y = "Proportion") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

ggplot(predictions, aes(x = age_range, y = success_probability, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Age",
       y = "Success Probability") +
  facet_grid(sex ~ .) +
  theme_minimal()

```
@fig-age-sex shows the predicted success probability according the the age and sex of the expedition member. We notice that the success probabilities over all ages tend to be slightly higher in males compared to females, just as we saw in @fig-model-coefficients. But, when combined with age we see there differences are not constant. We see that the difference in the success probabilities for people under 50 are very close to each other but for people aged 51 or above the difference is much larger. The success probabilities are pretty high, crossing 0.8 for younger individuals but starts declining with increase in age. The steep decline in category 81-90 might stem from the significant less number of expedition members in that age group leading to a less varied result.

@@

ggplot(predictions, aes(x = height_range, y = success_probability, fill = seasons)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Height of the Peak",
       y = "Success Probability") +
  facet_grid(seasons ~ .) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
@fig-height-seasons shows the predicted success probability according the the height of the peak being ascended and the season of expedition. We notice that the success probabilities are pretty high for shorter mountain ranges and become lesser with increase in height. This might have to do with the fact that less higher peaks could be comparatively easier to summit. This, however, cannot be established as the trend as we see a non-pattern being followed for heights 6800 metres and above. We see that winter season expeditions have lower success probabilities, probably owning to the extreme cold conditions at higher altitudes. We also notice interesting trend as an interaction of these two factors. One such trend would be the success probabilities during different seasons for peaks between 5400 metres and 5749 metres are all comparatively closer to each other than compared to peaks between 7500 metres and 7849 metres in height. All seasons' success probabilities for peaks between 5400 metres and 5749 metres fall between 0.85 to 0.9 whereas for for peaks between 7500 metres and 7848 metres, the success probabilities go from around 0.35 in Winter to around 0.6 in Autumn or Spring.

>>>>

plot_bar(analysis_data, "sex", "Sex")
plot_bar(analysis_data, "solo", "Solo Status")
plot_bar(analysis_data, "age_range", "Age")
plot_bar(analysis_data, "seasons", "Season")
plot_bar(analysis_data, "height_range", "Height of the Peak")
```

@@

  p <- ggplot(plot_data, aes(x = x_var, y = proportion, fill = factor(success))) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(
      values = c("TRUE" = "green", "FALSE" = "red"),
      labels = c("Success", "Died"),
      name = "Outcome"
    ) +
    labs(x = x_label, y = "Proportion") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

ggplot(predictions, aes(x = age_range, y = success_probability, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Age",
    y = "Success Probability"
  ) +
  facet_grid(sex ~ .) +
  theme_minimal()
```
@fig-age-sex shows the predicted success probability according the the age and sex of the expedition member. We notice that the success probabilities over all ages tend to be slightly higher in males compared to females, just as we saw in @fig-model-coefficients. But, when combined with age we see there differences are not constant. We see that the difference in the success probabilities for people under 50 are very close to each other but for people aged 51 or above the difference is much larger. The success probabilities are pretty high, crossing 0.8 for younger individuals but starts declining with increase in age. The steep decline in category 81-90 might stem from the significant less number of expedition members in that age group leading to a less varied result.

@@

ggplot(predictions, aes(x = height_range, y = success_probability, fill = seasons)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Height of the Peak",
    y = "Success Probability"
  ) +
  facet_grid(seasons ~ .) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
@fig-height-seasons shows the predicted success probability according the the height of the peak being ascended and the season of expedition. We notice that the success probabilities are pretty high for shorter mountain ranges and become lesser with increase in height. This might have to do with the fact that less higher peaks could be comparatively easier to summit. This, however, cannot be established as the trend as we see a non-pattern being followed for heights 6800 metres and above. We see that winter season expeditions have lower success probabilities, probably owning to the extreme cold conditions at higher altitudes. We also notice interesting trend as an interaction of these two factors. One such trend would be the success probabilities during different seasons for peaks between 5400 metres and 5749 metres are all comparatively closer to each other than compared to peaks between 7500 metres and 7849 metres in height. All seasons' success probabilities for peaks between 5400 metres and 5749 metres fall between 0.85 to 0.9 whereas for for peaks between 7500 metres and 7848 metres, the success probabilities go from around 0.35 in Winter to around 0.6 in Autumn or Spring.
